---
title: "Final Project Data Processing"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Load Library and reading the cleaned census data, where we only included 4 columns: AGEGRP: age (in groups); LFACT: labor force detail; YRIM: year of Immigration; IMMSTAT: immigrant status. These are the main columns that we will be performing our analysis on and try to explore the causal relationship. 
```{r}
rm(list=ls())

library(tidyverse)
library(haven)
library(stargazer)
library(ggplot2)
library(dplyr)

#Data Subsetting (previous preparatory work, commented out here)
#df_basic_d_y <- df |>
  #select(AGEGRP, LFACT, YRIM, IMMSTAT)

#head(df_basic_d_y)
#library(readr)

# write to a file called “df_basic_d_y.csv” in your working directory
#write_csv(df_basic_d_y, "cen_2021_clean.csv")

#Read in the cleaned dataset:
df <- read_csv("data/cen_2021_clean.csv")
head(df)
```

Data cleaning Process:
1) coerce all the PUMF code‐vars to numeric in one go
2) drop the “never worked” & missing codes in LFACT
3) keep only citizens or immigrants
4) restrict to the age groups you care about (AGEGRP 13–19)
5) create the Retired outcome variable (Yi)

```{r}
# df without useless columns (at the moment)
df_clean <- df %>%
  mutate(across(
    c(AGEGRP, LFACT, IMMSTAT, YRIM),
    ~ as.numeric(.)
  )) %>%
  
  filter(! LFACT %in% c(14, 88, 99)) %>%
  
  filter(IMMSTAT %in% c(1, 2)) %>%
  
  filter(AGEGRP >= 13, AGEGRP <= 19) %>%
  
  mutate(
    Retired = if_else(LFACT %in% c(11, 12, 13), 1L, 0L)
  )
head(df_clean)
```

# Quick check:
```{r}
df_clean %>% 
  count(LFACT)     # should only see 1–13
df_clean %>% 
  count(IMMSTAT)   # only 1,2
df_clean %>% 
  count(AGEGRP)    # only 13–19
df_clean %>% 
  count(Retired)   # retired vs. not
```

Separating treatment/ control group
```{r}
df_did <- df_clean %>%
  # define group membership once and for all
  mutate(
    eligible = as.integer(
      IMMSTAT == 1                          # all citizens
      | (IMMSTAT == 2 & YRIM <= 2011)      # immigrants ≥10 yrs
    ),
    post = as.integer(AGEGRP > 16),        # age 65+ if AGEGRP 17+
    did  = eligible * post                 # the actual treatment
  )
```

summary statistics
```{r}
summary(df_did)
```

```{r}
# 2. Summarise retirement rates by age and group
df_trends <- df_did %>%
  group_by(AGEGRP, eligible) %>%
  summarise(
    prop_retire = mean(Retired),
    .groups = "drop"
  )

```

```{r}
#Chat-GPT helped with the plotting partially
ggplot(df_trends, aes(x = AGEGRP, y = prop_retire, color = factor(eligible))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = 16.8, linetype = "dashed", color = "grey50") +
  scale_color_manual(
    values = c("blue", "red"),
    labels = c("Ineligible (control)", "Eligible (treatment)"),
    name   = "OAS Eligibility"
  ) +
  scale_x_continuous(breaks = 13:19) +
  labs(
    title    = "Retirement Rates by Age Group and OAS Eligibility",
    subtitle = "Full series (13–16 = pre-trend; jump at 17=65)",
    x        = "AGEGRP (13=50–54, …, 17=65–69, …)",
    y        = "Proportion Retired"
  ) +
  theme_minimal()
```

```{r}
# CTA-Restrict to pre‐treatment period: AGEGRP ≤ 16 (i.e. all below age‐65)
pre_df <- df_did %>% 
  filter(AGEGRP <= 16)

# Factor‐by‐factor approach
ct_fac <- lm(Retired ~ eligible * factor(AGEGRP), data = pre_df)
summary(ct_fac)

```

immigrant only: compare between treatment immigrants and control immigrants
```{r}
df_imm_only <- df_clean %>%
  # get rid of citizen
  filter(
    IMMSTAT == 2
  ) %>%
  # define group membership once and for all
  mutate(
    eligible = as.integer(YRIM <= 2011     # immigrants ≥10 yrs
    ),
    post = as.integer(AGEGRP > 16),        # age 65+ if AGEGRP 17+
    did  = eligible * post                 # the actual treatment
  )
```

```{r}
summary(df_imm_only)
```

```{r}
# 2. Summarise retirement rates by age and group
df_trends_imm <- df_imm_only %>%
  group_by(AGEGRP, eligible) %>%
  summarise(
    prop_retire = mean(Retired),
    .groups = "drop"
  )
```

```{r}
#Chat-GPT helped with the plotting partially
ggplot(df_trends_imm, aes(x = AGEGRP, y = prop_retire, color = factor(eligible))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = 16.8, linetype = "dashed", color = "grey50") +
  scale_color_manual(
    values = c("blue", "red"),
    labels = c("Ineligible (control)", "Eligible (treatment)"),
    name   = "OAS Eligibility"
  ) +
  scale_x_continuous(breaks = 13:19) +
  labs(
    title    = "Retirement Rates by Age Group and OAS Eligibility",
    subtitle = "Full series (13–16 = pre-trend; jump at 17=65)",
    x        = "AGEGRP (13=50–54, …, 17=65–69, …)",
    y        = "Proportion Retired"
  ) +
  theme_minimal()
```
```{r}
# CTA-Restrict to pre‐treatment period(imm only): AGEGRP ≤ 16 (i.e. all below age‐65)
pre_df_imm <- df_imm_only %>% 
  filter(AGEGRP <= 16)

# Factor‐by‐factor approach
ct_fac <- lm(Retired ~ eligible * factor(AGEGRP), data = pre_df_imm)
summary(ct_fac)
```

## Regression models

1. baseline (additive) model for citizen-immigrant dataset:
```{r}
lm_baseline_c_i <- lm(Retired~ post + eligible, data=df_did)

summary(lm_baseline_c_i)
```

2.1 did linear probability model for citizen-immigrant dataset:
```{r}
lm_did_c_i <- lm(Retired~ post * eligible, data=df_did)

summary(lm_did_c_i)
```
2.2 did logistic model for citizen-immigrant dataset:
```{r}
logm_did_c_i <- glm(Retired~ post * eligible, data=df_did, family=binomial)

summary(logm_did_c_i)
```

3. baseline (additive) model for immigrant-only sample:
```{r}
lm_baseline_i <- lm(Retired~ post + eligible, data=df_imm_only)

summary(lm_baseline_i)
```
4.1 did linear probability model for immigrant-only sample:
```{r}
lm_did_i <- lm(Retired~ post * eligible, data=df_imm_only)

summary(lm_did_i)
```
4.2 did logistic model for immigrant-only sample:
```{r}
logm_did_i <- glm(Retired~ post * eligible, data=df_imm_only, family=binomial)

summary(logm_did_i)
```
Citizen–Immigrant Models
```{r}
#Code generated under help of Chat-GPT
stargazer(
  lm_baseline_c_i, lm_did_c_i, logm_did_c_i,
  type = "html",
  out = "table_citizen_immigrant.html",
  title = "Citizen-Immigrant Models: OAS Elibility Effects",
  column.labels = c("Baseline", "DID (LPM)", "DID (Logit)"),
  covariate.labels = c("Post", "Eligible", "Post × Eligible"),
  dep.var.labels = "Retired",
  omit.stat = c("f", "ser"),
  report = "vc*",
  digits = 3,
  no.space = TRUE,
  align = TRUE,
  notes = "Standard errors in parentheses. *** p<0.01, ** p<0.05, * p<0.1"
)


```

 Immigrant-Only Models
```{r}
#Code generated under help of Chat-GPT
stargazer(
  lm_baseline_i, lm_did_i, logm_did_i,
  type = "html",
  out = "table_immigrant_only.html",
  title = "Immigrant-Only Models: OAS Elibility Effects",
  column.labels = c("Baseline", "DID (LPM)", "DID (Logit)"),
  covariate.labels = c("Post", "Eligible", "Post × Eligible"),
  dep.var.labels = "Retired",
  omit.stat = c("f", "ser"),
  report = "vc*",
  digits = 3,
  no.space = TRUE,
  align = TRUE,
  notes = "Standard errors in parentheses. *** p<0.01, ** p<0.05, * p<0.1"
)


```
